Estimating Access to Rabies Post-Exposure Prophylaxis in Madagascar

### 2018-12-04 accessPEP.Rmd
district models

## poisson with total pop as offset
mod <- glm(round(bites) ~ ttimes_weighted + pop + offset(log(round(pop))), 
                 data = exps_cleaned, family = poisson)
plot(exps_cleaned$bites, predict(mod, type = "response"))
plot(exps_cleaned$ttimes_weighted, predict(mod, type = "response"))

@@

```{r}
p_rabid = 0.50
mean_bites <- predict(mod, type = "response", newdata = covars)
bites <- rpois(length(mean_bites), mean_bites)
mean_exps <- runif(length(covars$pop), min = 0.39*0.01*exps_cleaned$pop/25, 
                    max = 0.39*0.01*exps_cleaned$pop/5)
rabid_exps <- rpois(length(mean_bites), mean_exps)
p_rabid_t <- ifelse(bites*p_rabid/rabid_exps > 0.95, (0.95*rabid_exps)/bites, p_rabid)
deaths <- rbinom(length(mean_bites), rabid_exps - round(bites*p_rabid_t), 0.16)
rho <- bites*p_rabid_t/rabid_exps
plot(covars$ttimes_weighted, rho)

@@

@@

quartz()
ggplot() +
  geom_polygon(data = mada_dist_gg, aes(x = long, y = lat, group = group, fill = CTAR)) +
  scale_colour_viridis_d()

>>>>

## poisson with total pop as offset
mod <- glm(round(bites) ~ ttimes_weighted + offset(log(round(pop))), 
                 data = exps_cleaned, family = poisson)
plot(exps_cleaned$bites, predict(mod, type = "response"))
plot(exps_cleaned$ttimes_weighted, predict(mod, type = "response"))

@@

```{r}
get.burden <- function(covs = covars, pop = covars$pop, model = mod,
                       p_rabid = 0.5, rho_max = 0.9,
                       max_HDR = 25, min_HDR = 5, 
                       dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                       prob_death = 0.16, nsims = 1000) {
  store_deaths <- store_p_rabid <- matrix(NA, nrow(covs), nsims)

  for (i in 1:nsims) {
    mean_bites <- predict(model, type = "response", newdata = covs)
    bites <- rpois(length(mean_bites), mean_bites)
    mean_exps <- runif(length(pop), min = human_exp_rate*dog_rabies_inc*pop/max_HDR, 
                    max = human_exp_rate*dog_rabies_inc*pop/min_HDR)
    rabid_exps <- rpois(length(mean_bites), mean_exps)
    p_rabid_t <- ifelse(bites*p_rabid/rabid_exps > rho_max, (rho_max*rabid_exps)/bites, p_rabid)
    deaths <- rbinom(length(mean_bites), rabid_exps - round(bites*p_rabid_t), prob_death)
    store_deaths[, i] <- deaths
    store_p_rabid[, i] <- p_rabid_t
  }
  results <- cbind(get.meanCI(store_deaths), get.meanCI(store_p_rabid))
  colnames(results) <- c("deaths_mean", "deaths_lowerCI", "deaths_upperCI",
                         "p_rabid_mean", "p_rabid_lowerCI", "p_rabid_upperCI")
  results <- cbind(covars, results)
  return(results)
}

get.burden.flat <- function(covs = covars, pop = covars$pop, model = mod,
                       p_rabid = 0.5, rho = 0.85, 
                       prob_death = 0.16, nsims = 1000) {
  store_deaths <- matrix(NA, nrow(covs), nsims)

  for (i in 1:nsims) {
    mean_bites <- predict(model, type = "response", newdata = covs)
    bites <- rpois(length(mean_bites), mean_bites)
    unreported <- bites*p_rabid/rho - bites*p_rabid
    deaths <- rbinom(length(mean_bites), round(unreported), prob_death)
    store_deaths[, i] <- deaths
  }
  results <- get.meanCI(store_deaths)
  colnames(results) <- c("deaths_mean", "death_lowerCI", "deaths_upperCI")
  results <- cbind(covars, results)
  return(results)
}

get.meanCI <- function(matrix_by_row, nboots = 1000) {
    mean_rows <- apply(matrix_by_row, 1, mean)
    CIs <- apply(matrix_by_row, 1, boots, nboot = nboots)
    lower <- unlist(CIs)[1, ] 
    upper <- unlist(CIs)[2, ]
    summ <- cbind(mean_rows, lower, upper)
    colnames(summ) <- c("mean", "lower", "upper")
    return(summ)
}

boots <- function(x, nboot = 1000){
  n = length(x)
  xbar = mean(x)
  tmpdata = sample(x,n*nboot, replace=TRUE)
  bootstrapsample = matrix(tmpdata, nrow=n, ncol=nboot)
  bsmeans = colMeans(bootstrapsample)
  deltastar = bsmeans - xbar
  d = quantile(deltastar, c(0.05, 0.95))
  ## Get from quantiles
  ci = xbar - c(d[2], d[1])

  ## Sorted CIs
  sorteddeltastar = sort(deltastar)
  # # Look at the sorted results
  # Find the .05 and .95 critical values of deltastar
  d9alt = sorteddeltastar[5]
  d1alt = sorteddeltastar[950]
  # Find and print the 80% confidence interval for the mean
  ciAlt = xbar - c(d1alt,d9alt)
  # cat(’Alternative confidence interval: ’,ciAlt, ’\n’)
  return(ci)
}

@@

```{r}
results_pr20 <- get.burden(p_rabid = 0.20, rho_max = 0.98)
results_pr60 <- get.burden(p_rabid = 0.60,rho_max = 0.98)

plot(results_pr20$ttimes_weighted/60, results_pr20$deaths_mean, col = "darkred", pch = 20)
points(results_pr60$ttimes_weighted/60, results_pr60$deaths_mean, col = "red", pch = 1)
segments(x0 = results_pr60$ttimes_weighted/60, x1 = results_pr60$ttimes_weighted/60,
         y0 = results_pr60$deaths_mean, y1 = results_pr20$deaths_mean, col = "red")

results_pr20 <- get.burden(p_rabid = 0.20, rho_max = 0.90)
results_pr60 <- get.burden(p_rabid = 0.20, rho_max = 0.99)

plot(results_pr20$ttimes_weighted/60, results_pr20$deaths_mean, col = "darkred", pch = 20)
points(results_pr60$ttimes_weighted/60, results_pr60$deaths_mean, col = "red", pch = 1)
segments(x0 = results_pr60$ttimes_weighted/60, x1 = results_pr60$ttimes_weighted/60,
         y0 = results_pr60$deaths_mean, y1 = results_pr20$deaths_mean, col = "red")

results_pr20 <- get.burden(p_rabid = 0.60, rho_max = 0.90)
results_pr60 <- get.burden(p_rabid = 0.60, rho_max = 0.99)

plot(results_pr20$ttimes_weighted/60, results_pr20$deaths_mean, col = "darkred", pch = 20)
points(results_pr60$ttimes_weighted/60, results_pr60$deaths_mean, col = "red", pch = 1)
segments(x0 = results_pr60$ttimes_weighted/60, x1 = results_pr60$ttimes_weighted/60,
         y0 = results_pr60$deaths_mean, y1 = results_pr20$deaths_mean, col = "red")
```

@@

ggplot() +
  geom_polygon(data = mada_dist_gg, aes(x = long, y = lat, group = group, fill = CTAR)) +
  scale_colour_viridis_d()
